The Revisited Hidden Weight Bit Function

### 2024-11-28 RevisedHiddenWeight.tex
- Corrections, plus indications for parts which need to be addressed.

	\titlerunning{The Revisited Hidden Weight Bit Function}
\author{
	Pierrick M\'eaux\inst{1},
	Tim Seuré\inst{1}, 
	Deng Tang\inst{2}
%	Pierrick M\'eaux\orcidID{0000-0001-5733-4341}\inst{1},
%	Tim Seuré\inst{1}, 
%	Deng Tang\orcidID{0000-0002-8373-9200}\inst{2}
}
\authorrunning{P. M\'eaux, T. Seuré, D. Tang}

@@

@@

Recent developments in stream ciphers have reignited interest in this function. The emergence of new applications for stream ciphers with filter functions on a larger number of variables—such as Hybrid Homomorphic Encryption (\textsf{HHE})~\cite{CCS:NaeLauVai11}—has further emphasized the relevance of such functions. \textsf{HHE} requires Boolean functions that can be efficiently evaluated in an input-oblivious algorithm. For instance, several new binary stream ciphers have been proposed since 2016, including Kreyvium~\cite{JOC:CCFLNP18}, FLIP~\cite{EC:MJSC16}, Rasta~\cite{C:Rasta}, FiLIP~\cite{INDO:MCJS19}, Dasta~\cite{TOSC:BeiLea20} and Fasta~\cite{CTRSA:CidIndRad22}. 
Currently, designs employing Boolean functions with more than one hundred variables as filters lead to \textsf{HHE} schemes with the best latency~\cite{INDO:HofMeaRic20,CCS:CDPP22,CIC:MeaParPei24,eprint:AGHM24}.

@@

Throughout this paper, $n$ will always denote a positive integer. Further, our intervals will only contain integers, so that for instance $[0,n]=\{0,1,\dots,n\}$. The set of binary vectors of length $n$ will be denoted by $\F_2^n$. The entries of a binary vector $x\in\F_2^n$ will always be denoted by $x_1,\dots,x_n$. We denote the scalar product of two binary vectors $x,y\in\F_2^n$ by $x\cdot y=\sum_{i=1}^nx_iy_i\in\F_2$. The Hamming weight of a binary vector $x\in\F_2^n$ is denoted by $\w(x)=|\{i\in [1,n]:x_i=1\}|$.

@@

\begin{Prop}[Properties of some functions from $\WD{1}{n}$\label{prop:wwd1}]
	Let $n\geq 2$ be an even integer.
	\begin{itemize}
		\item (\eg{}~\cite{DCC:DalMaiSar06}, Theorem $3$) The majority function $\mathsf{Maj}_n$ introduced in Definition~\ref{def:maj} satisfies:
		\[
		\NL(\mathsf{Maj}_n)=2^{n-1}-\binom{n-1}{\frac n2}.
		\]
		\item (\cite{DAM:WCST14}, Theorem $3$) The \hwbf{} $h$ introduced in Definition~\ref{def:hwbf} satisfies:
		\[
		\NL(h)=2^{n-1}-2\binom{n-2}{\frac{n-2}{2}}.
		\]	
		\item (\cite{DAM:WCST14}, Theorem $1$) The \hwbf{} $h$ introduced in Definition~\ref{def:hwbf} satisfies $\degg(h)=n-1$ for all $n\geq 4$.
	\end{itemize}
\end{Prop}

\subsection{Krawtchouk polynomials}
We use Krawtchouk polynomials and some of their properties to prove one of our main results. We give the necessary preliminaries here and refer to \eg{}~\cite{book:MacSlo78} for more details.

\begin{definition}[Krawtchouk polynomials]\label{def:Kraw}
	The $n$-th Krawtchouk polynomial of degree $k\in[0,n]$ is given by:
    \[
        \displaystyle \kraw{k}{x}{n}=\sum_{j=0}^{k} (-1)^j \binom{x}{j} \binom{n-x}{k-j}.
    \]
    Alternatively, the Krawtchouk polynomials can be characterized as the coefficients in the generating function for $(1+z)^{n-x} (1-z)^x$ in the variable $z$:
    \[
        (1+z)^{n-x} (1-z)^x=\sum_{k=0}^\infty \kraw{k}{x}{n} z^k.
    \]
\end{definition}

@@

\begin{definition}[Revisited \hwbf{}]\label{def:revHWBF}
Let $n\in \N^*$, even, we call revisited \hwbf{} and denote by $f$ the Boolean function defined as:
\[f(x_1,\cdots,x_n)=\left(\sum_{i=1}^{n/2} (x_i+1) x_{i+\frac{n}{2}}\right) + \sum_{k=1}^n \phikn{k}{n} x_k.\]
\end{definition}

%We study the function $f$ in an even number of variables $n$ defined as the following:
%\[f(x_1,\cdots,x_n)=\left(\sum_{i=1}^{n/2} (x_i+1) x_{i+\frac{n}{2}}\right) + \sum_{k=1}^n \phikn{k}{n} x_k.\]

Since $f$ is the sum of a quadratic function and a weightwise affine function, it is a weightwise quadratic function. 
Using the formalism of Definition~\ref{def:wwdegd}, $f$ is the weightwise quadratic function defined by:% $f_0=0$ and for $k\in [1,n]$ $f_k=x_k + \sum_{i=1}^{n/2} (x_i+1) x_{i+n/2}$.
\[f_0=0, \quad \text{ and } \quad \forall k \in [1,n], \, f_k=x_k+\sum_{i=1}^{n/2} (x_i+1) x_{i+\frac{n}{2}}.\]

This form allows to derive the restricted Walsh transform of $f$, which can be a useful tool to study the balancedness and bound the nonlinearity of a function.
%Maybe add examples, such as OM24, M24

%We study the restricted Walsh transform  of $f$ to bound its nonlinearity.
\begin{proposition}\label{prop:restrWT}
Let $n \in \N^*$ even, and $f$ be the function from Definition~\ref{def:revHWBF}, for all $k\in [0,n]$ its restricted Walsh transform in $a\in \F_2^n$ is given by:
\[\wtk{f}{k}(a)=\sum_{x \in \Ekn{k}{n}} (-1)^{\sum_{i=1}^{n/2} x_i x_{i+\frac{n}{2}}+ c \cdot x},\]
where $c$ is the $n$-length binary vector such that $c=a+e_k+\sum_{i=n/2+1}^n e_i$ where the $e_i$ denotes the canonical vector with $1$ in position $i$.
\end{proposition}
\begin{proof}
%\begin{align*}
%	\wtk{f}{k}(a)&=\sum_{x \in \Ekn{k}{n}} (-1)^{f(x)+a \cdot x}
%	=\sum_{x \in \Ekn{k}{n}} (-1)^{x_k + \sum_{i=1}^{n/2} (x_i+1) x_{i+\frac{n}{2}}+ a \cdot x}\\
%	&=\sum_{x \in \Ekn{k}{n}} (-1)^{\sum_{i=1}^{n/2} x_i x_{i+\frac{n}{2}}+ c \cdot x},
%\end{align*}
%\begin{align*}
\[	\wtk{f}{k}(a)=\sum_{x \in \Ekn{k}{n}} (-1)^{f(x)+a \cdot x}
	=\sum_{x \in \Ekn{k}{n}} (-1)^{x_k + \sum_{i=1}^{n/2} (x_i+1) x_{i+\frac{n}{2}}+ a \cdot x}\\
	=\sum_{x \in \Ekn{k}{n}} (-1)^{\sum_{i=1}^{n/2} x_i x_{i+\frac{n}{2}}+ c \cdot x},\]
where $c$ denotes the $n$-length binary vector such that $c=a+e_k+\sum_{i=n/2+1}^n e_i$ and $e_i$ denotes the $i$-th canonical vector.
\end{proof}

@@

Accordingly, we can study the restricted Walsh transform of this function by analyzing the one of $\sum_{i=1}^{n/2} x_i x_{i+\frac{n}{2}}$, for all $a$. We introduce the following notations.

\begin{definition}
Let $n\in \N$ be an even number, and $a\in \F_2^n$. We denote $d_n$ the function given by: $d_n(x)=\sum_{i \text{ odd}}^n x_i x_{i+1}$, 
and we denote by $\Dnka{n}{k}{a}$ the quantity:
\[\Dnka{n}{k}{a}= \sum_{x\in \Ekn{k}{n}} (-1)^{d_n(x) +a x}\]
\end{definition}

We can give a recursive relation for $\Dnka{n}{k}{a}$.

\begin{proposition}\label{prop:recursiveDnka}
Let $n\in \N$ be an even number, $k$ an integer in the range $[0,n]$ and $a\in \F_2^n$. the Following holds on $\Dnka{n}{k}{a}$:
\begin{itemize}
	\item For $n=0$, $\Dnka{0}{k}{a}=1$ if $k=0$ and $\Dnka 0ka=0$ otherwise.
	\item if $n\geq 2$, denoting  $b=(a_1,\cdots,a_{n-2})$ and $y=(x_1,\cdots,x_{n-2})$, we get:
	\[\Dnka{n}{k}{a}=\Dnka{n-2}{k}{b}+\left((-1)^{a_{n-1}} + (-1)^{a_{n}}\right)\Dnka{n-2}{k-1}{b} + (-1)^{1+a_{n-1}+a_n} \Dnka{n-2}{k-2}{b}.\]
\end{itemize}

\begin{proof}
For $n\geq 2$ we can rewrite $\Dnka{n}{k}{a}$:

\begin{align*}
\Dnka{n}{k}{a}&=\sum_{x\in \Ekn{k}{n}} (-1)^{d_n(x) +a x}\\
&=\sum_{x\in \Ekn{k}{n-2}} (-1)^{d_{n-2}(y) +b y}+
\sum_{x\in \Ekn{k-1}{n-2}} (-1)^{d_{n-2}(y) +b y + a_{n-1}}\\
&+\sum_{x\in \Ekn{k-1}{n-2}} (-1)^{d_{n-2}(y) +b y + a_{n}}
+\sum_{x\in \Ekn{k-2}{n-2}} (-1)^{d_{n-2}(y) +b y +1 + a_{n-1}+ a_{n}}\\
&=\Dnka{n-2}{k}{b}+\left((-1)^{a_{n-1}} + (-1)^{a_{n}}\right)\Dnka{n-2}{k-1}{b} + (-1)^{1+a_{n-1}+a_n} \Dnka{n-2}{k-2}{b}.
\end{align*}

The limit cases $\Dnka 0ka=1$ if $k=0$ and $\Dnka 0ka=0$ otherwise are obtained from the values of $\Dnka{2}{k}{a}$ given in Table~\ref{tab:Dnka}.
\end{proof}

@@

The recursive formula of Proposition~\ref{prop:recursiveDnka} gives different cases depending on the values of $a_{n-1}$ and $a_n$, however the same reasoning applies with any $a_i, a_{i+1}$ where $i$ is odd. 
Thereafter, the value of $\Dnka{n}{k}{a}$ depends only on the number of pairs of $a$ being $(0,0)$, $(1,1)$ or $(0,1)$ or $(1,0)$, and the values $\Dnka{2}{k}{a}$ for $k \in [0,2]$ and $a \in \{ (0,0),(1,1),(0,1)\}$. 
We give the first values of $\Dnka{2}{k}{a}$ in Table~\ref{tab:Dnka},  these values and Proposition~\ref{prop:recursiveDnka} are sufficient to determine any $\Dnka{n}{k}{a}$. 
In the following, we show how the Walsh transform of $f$ can be written in terms of $\Dnka{n}{k}{a}$. Then, in Section~\ref{sec:balancedness} we use Proposition~\ref{prop:recursiveDnka} to determine the balancedness of $f$, and in Section~\ref{sec:genfunc} we study the value of $\Dnka{n}{k}{a}$ using trough generating functions.

%The first values of $\Dnka{2}{k}{a}$ are listed in Table~\ref{tab:Dnka}

\begin{table}[H]
	\centering
	\begin{tabular}{|c|c|c|c|}
		\hline
		$a$ & $(0,0)$ & $(0,1)$ & $(1,1)$ \\ \hline
		$\Dnka{2}{0}{a}$ &$1$&$1$ & $1$\\ \hline
		$\Dnka{2}{1}{a}$ &$2$& $0$ & $-2$\\ \hline
		$\Dnka{2}{2}{a}$ &$-1$&$1$ & $-1$\\ \hline
	\end{tabular}
	\tablecaption{Values of $\Dnka{2}{k}{a}$}\label{tab:Dnka}
\end{table}

Then, we can express the restricted Walsh transform of $f$ in terms of $\Dnka{n}{k}{a}$:

\begin{proposition}[Walsh transform of $f$]\label{prop:WT}
	Let $n \in \N^*$ even, and $f$ be the function from Definition~\ref{def:revHWBF}, for all $a\in \F_2^n$ its Walsh transform is given by:
	\[ \wt{f}(a)=1+\sum_{k=1}^n \Dnka{n}{k}{b+\pi(e_k)},\]
where $\pi$ is the permutation sending the first $n/2$ elements to the odd positions, and the $n/2$ last ones to the even positions, and $b=\pi(a)+(0,1,0,1,\ldots,0,1)$.
\end{proposition}

\begin{proof}
Using Proposition~\ref{prop:restrWT}, the restricted Walsh transform of $f$ can then be written as \[
\wtk{f}{k}(a)=\sum_{x \in \Ekn{k}{n}} (-1)^{f(x)+ax}=\sum_{x \in \Ekn{k}{n}} (-1)^{\sum_{i=1}^{n/2} x_i x_{i+\frac{n}{2}}+ c \cdot x}=\Dnka{n}{k}{d},\]
where $d$ is the permutation of $c$ obtained by sending the first $n/2$ elements to the odd positions, and the $n/2$ last ones to the even positions.

@@

\subsection{Balancedness of $f$}\label{sec:balancedness}

@@

\begin{theorem}[Balancedness of $f$]\label{th:bal}
	Let $n \in \N^*$ even, and $f$ be the $n$-variable function from Definition~\ref{def:revHWBF}, the following holds on its Walsh transform in $0_n$:
	\[\wt{f}(0_n)= \left \{
	\begin{array}{l l}
	0 & \text{ if } n/2 \text{ is even, } \\
	-2 \binom{n/2-1}{(n-2)/4} & \text{ if }  n/2 \text{ is odd.}
	\end{array}\right.\]

@@

In this section, we study the values of $\Dnka nka$ using generating functions. First, in Section~\ref{sec:genfunc} we establish key properties of these quantities. Then, in Section~\ref{sec:Cauchy}, we apply Cauchy's estimate to bound the absolute values of $\Dnka nka$, which enables us to bound the absolute values of the Walsh transform coefficients of $f$ and consequently, the nonlinearity.
Finally, in Section~\ref{sec:general}, we demonstrate how this result can be extended to bound the nonlinearity of a family of weightwise quadratic functions.

\subsection{Study of $\Dnka nka$ through Generating Functions}\label{sec:genfunc}

\begin{definition}\label{defi:P_a}
    For $a\in\F_2^n$, denote by $p=p(a)$ the number of odd $i\geq 1$ for which $(a_i,a_{i+1})=(0,0)$, by $q=q(a)$ the number of odd $i\geq 1$ for which $(a_i,a_{i+1})=(1,1)$, and by $r=r(a)$ the number of odd $i\geq 1$ for which $(a_i,a_{i+1})\in\{(0,1),(1,0)\}$. We then have $p+q+r=n/2$. Consider further the integer polynomial $P_a(z)$ given by the following expression:
    \[
    P_a(z)=(-z^2+2z+1)^{p}\cdot(-z^2-2z+1)^{q}\cdot(z^2+1)^{r}.
    \]
\end{definition}

\begin{proposition}\label{proposition:generating_fct}
	It holds that $\sum_{k\in\Z}\Dnka nka z^k= P_a(z)$.
\end{proposition}

\begin{proof}
	We proceed by induction on even $n\geq 0$. 
	For the base case $n=0$, for $a\in\F_2^0$, we have $P_a(z)=1$, while $\Dnka 0ka=1$ if $k=0$ and $\Dnka 0ka=0$ otherwise (from Proposition~\ref{prop:recursiveDnka}). 
	For the inductive step, let $n\geq 2$ be even, and assume that the corresponding formula holds for $n-2$. Using the recursive formula for $\Dnka nka$ with the notation $a=(b,a_{n-1},a_n)$, we have:
	\begin{align*}
	\begin{split}
	\sum_{k\in\Z}\Dnka nka z^k&=\sum_{k\in\Z}\Dnka{n-2}kbz^k+\big((-1)^{a_{n-1}}+(-1)^{a_{n}}\big)\sum_{k\in\Z}\Dnka{n-2}{k-1}{b}z^k\\
	&\quad+(-1)^{1+a_{n-1}+a_n}\sum_{k\in\Z}\Dnka{n-2}{k-2}{b}z^k\\
	&=P_b(z)+\big((-1)^{a_{n-1}}+(-1)^{a_{n}}\big)zP_b(z)+(-1)^{1+a_{n-1}+a_n}z^2P_b(z)\\
	&=P_b(z)\cdot\Big(z^2(-1)^{1+a_{n-1}+a_n}+z\big((-1)^{a_{n-1}}+(-1)^{a_{n}}\big)+1\Big)\\
	&=P_a(z).\qedhere
	\end{split}
	\end{align*}
\end{proof}

\begin{remark}
    In particular, since $P_a(z)$ only depends on the triplet $(p,q,r)$, we deduce that also $\Dnka nka$ only depends on the triplet $(p,q,r)$; therefore, it makes sense to introduce the notation $\mathsf D_{n,k}^{p,q,r}=\Dnka nka$, where $a\in\mathbb F_2^n$ is any vector with $(p(a),q(a),r(a))=(p,q,r)$. We will make use of this notation later on.
\end{remark}

\begin{proposition}\label{proposition:symmetry}
	For every integer $k$, we have:
	\[	\Dnka n{n-k}a=(-1)^{p+q+k}\Dnka nka.	\]
\end{proposition}

>>>>

	\titlerunning{The Revisited Hidden Weight Bit Function}
\author{
	Pierrick M\'eaux\inst{1},
	Tim Seuré\inst{1}, 
	Deng Tang\inst{2}
%	Pierrick M\'eaux\orcidID{0000-0001-5733-4341}\inst{1},
%	Tim Seuré\inst{1}, 
%	Deng Tang\orcidID{0000-0002-8373-9200}\inst{2}
}
\authorrunning{P.\ M\'eaux, T.\ Seur\'e, D.\ Tang}

@@

\tsnote{@Self, check entire document.}

@@

Recent developments in stream ciphers have reignited interest in this function. The emergence of new applications for stream ciphers with filter functions on a larger number of variables—such as Hybrid Homomorphic Encryption (\textsf{HHE})~\cite{CCS:NaeLauVai11}—has further emphasized the relevance of such functions. \textsf{HHE} requires Boolean functions that can be efficiently evaluated in an input-oblivious algorithm. For instance, several new binary stream ciphers have been proposed since 2016, including Kreyvium~\cite{JOC:CCFLNP18}, \textsf{FLIP}~\cite{EC:MJSC16}, Rasta~\cite{C:Rasta}, \textsf{FiLIP}~\cite{INDO:MCJS19}, Dasta~\cite{TOSC:BeiLea20} and Fasta~\cite{CTRSA:CidIndRad22}. 
Currently, designs employing Boolean functions with more than one hundred variables as filters lead to \textsf{HHE} schemes with the best latency~\cite{INDO:HofMeaRic20,CCS:CDPP22,CIC:MeaParPei24,eprint:AGHM24}.

@@

Throughout this paper, $n\geq 0$ will always denote a non-negative integer. Further, our intervals will only contain integers, so that for instance $[0,n]=\{0,1,\dots,n\}$. The set of binary vectors of length $n$ will be denoted by $\F_2^n$, with zero vector written as $0_n$. The canonical basis of $\F_2^n$ will be written as $\{e_1,\dots,e_n\}$; therefore, $e_i\in\F_2^n$ is the vector which is zero everywhere, except at position $i$. The entries of a binary vector $x\in\F_2^n$ will always be denoted by $x_1,\dots,x_n$. Given a permutation $\pi:[1,n]\to[1,n]$ and a vector $x\in\F_2^n$, we define $\pi(x)=(x_{\pi(i)})_{i\in[1,n]}\in\F_2^n$. We write the scalar product of two binary vectors $x,y\in\F_2^n$ as $x\cdot y=\sum_{i=1}^nx_iy_i\in\F_2$. The Hamming weight of a binary vector $x\in\F_2^n$ is denoted by $\w(x)=|\{i\in [1,n]:x_i=1\}|$.

@@

\begin{Prop}\label{prop:wwd1}
	Let $n\geq 2$ be an even integer.
	\begin{itemize}
		\item (\eg{} Theorem $3$ in~\cite{DCC:DalMaiSar06}) The majority function $\mathsf{Maj}_n$ introduced in Definition~\ref{def:maj} satisfies:
		\[
		\NL(\mathsf{Maj}_n)=2^{n-1}-\binom{n-1}{\frac n2}.
		\]
		\item (Theorem $3$ in~\cite{DAM:WCST14}) The \hwbf{} $h$ introduced in Definition~\ref{def:hwbf} satisfies:
		\[
		\NL(h)=2^{n-1}-2\binom{n-2}{\frac{n-2}{2}}.
		\]	
		\item (Theorem $1$ in \cite{DAM:WCST14}) The \hwbf{} $h$ introduced in Definition~\ref{def:hwbf} satisfies $\degg(h)=n-1$ for all $n\geq 4$.
	\end{itemize}
\end{Prop}

\subsection{Krawtchouk polynomials}
We use Krawtchouk polynomials and some of their properties to prove one of our main results. We give the necessary definition here and refer to \eg{}~\cite{book:MacSlo78} for more details.

\begin{definition}[Krawtchouk polynomials]\label{def:Kraw}
	The $n$-th Krawtchouk polynomial of degree $k\in[0,n]$ is given by:
    \[
        \kraw{k}{x}{n}=\sum_{j=0}^{k} (-1)^j \binom{x}{j} \binom{n-x}{k-j}.
    \]
    If $k<0$, we further define $\kraw{k}{x}{n}=0$. The Krawtchouk polynomials can alternatively be characterized as the coefficients in the generating function for $(1+z)^{n-x} (1-z)^x$ in the variable $z$:
    \[
        (1+z)^{n-x} (1-z)^x=\sum_{k\in\Z} \kraw{k}{x}{n} z^k.
    \]
\end{definition}

@@

\begin{definition}[Revisited \hwbf{}]\label{def:revHWBF}
For an even integer $n\geq 0$, we call revisited \hwbf{} the Boolean function $f\in\BN$ defined as:
\[f(x)=\sum_{i=1}^{n/2}(x_i+1) x_{i+\frac{n}{2}} + \sum_{k=1}^nx_k\phikn{k}{n}(x).\]
\end{definition}

Since $f$ is the sum of a quadratic function and a weightwise affine function, it is a weightwise quadratic function. Using the formalism of Definition~\ref{def:wwdegd}, $f$ is the weightwise quadratic function defined by $f_0(x)=0$ and $f_k(x)=x_k+\sum_{i=1}^{n/2}(x_i+1) x_{i+n/2}$ for $k\in[1,n]$. This form allows to derive the restricted Walsh transform of $f$, which can be a useful tool to study the balancedness and bound the nonlinearity of a function.
%Maybe add examples, such as OM24, M24
%We study the restricted Walsh transform  of $f$ to bound its nonlinearity.
\begin{proposition}\label{prop:restrWT}
Let $n\geq 0$ be an even integer, and let $f\in\BN$ be the revisited \hwbf{}. Consider any binary vector $a\in\F_2^n$, and define $c=a+e_k+\sum_{i=n/2+1}^n e_i$. Then the restricted Walsh transform of $f$ satisfies:
\[\wtk{f}{k}(a)=\sum_{x \in \Ekn{k}{n}} (-1)^{\sum_{i=1}^{n/2} x_i x_{i+n/2}+ c \cdot x}.\]
\end{proposition}
\begin{proof}
We compute:
%\begin{align*}
%	\wtk{f}{k}(a)&=\sum_{x \in \Ekn{k}{n}} (-1)^{f(x)+a \cdot x}
%	=\sum_{x \in \Ekn{k}{n}} (-1)^{x_k + \sum_{i=1}^{n/2} (x_i+1) x_{i+\frac{n}{2}}+ a \cdot x}\\
%	&=\sum_{x \in \Ekn{k}{n}} (-1)^{\sum_{i=1}^{n/2} x_i x_{i+\frac{n}{2}}+ c \cdot x},
%\end{align*}
%\begin{align*}
\begin{align*}
    \wtk{f}{k}(a)&=\sum_{x \in \Ekn{k}{n}} (-1)^{f(x)+a \cdot x}\\
	&=\sum_{x \in \Ekn{k}{n}} (-1)^{x_k + \sum_{i=1}^{n/2} (x_i+1) x_{i+n/2}+ a \cdot x}\\
	&=\sum_{x \in \Ekn{k}{n}} (-1)^{\sum_{i=1}^{n/2} x_i x_{i+n/2}+\left(x_k+\sum_{i=n/2+1}^{n}x_i+a\cdot x\right)}\\
    &=\sum_{x \in \Ekn{k}{n}} (-1)^{\sum_{i=1}^{n/2} x_i x_{i+n/2}+ c \cdot x}.\qedhere
\end{align*}
\end{proof}

@@

Accordingly, we can study the restricted Walsh transform of the revisited \hwbf{} by analyzing the one of the following Boolean function $d_n\in\BN$:

\begin{definition}
Let $n\geq 0$ be an even integer. Then we define the Boolean function $d_n\in\BN$ by $d_n(x)=\sum_{i=1}^{n/2}x_{2i-1}x_{2i}$. Further, for an integer $k$ and a binary vector $a\in\F_2^n$, we define $\Dnka{n}{k}{a}=0$ if $n\not\in[0,n]$, and otherwise:
\[
    \Dnka{n}{k}{a}=\sum_{x\in \Ekn{k}{n}} (-1)^{d_n(x) +a\cdot x}.
\]
\end{definition}

The $\Dnka{n}{k}{a}$ satisfy a recursive relation:

\begin{proposition}\label{prop:recursiveDnka}
Let $n\geq 0$ be an even integer. Then for all $k\in[0,n]$ and $a\in\F_2^n$, the following hold true:
\begin{itemize}
    \item We have $\Dnka 00a=1$ and $\Dnka 0ka=0$ if $k\neq 0$.
    \item If $n\geq 2$, then for $b=(a_1,\dots,a_{n-2})\in\F_2^{n-2}$, we have:
    \[\Dnka{n}{k}{a}=\Dnka{n-2}{k}{b}+\left((-1)^{a_{n-1}} + (-1)^{a_{n}}\right)\Dnka{n-2}{k-1}{b} + (-1)^{1+a_{n-1}+a_n} \Dnka{n-2}{k-2}{b}.\]
\end{itemize}
\end{proposition}

\begin{proof}
    We have that $\Dnka 00a=(-1)^0=0$ since $\Ekn 00=\{\epsilon\}$ for the empty vector $\epsilon\in\F_2^0$. That $\Dnka 0ka=0$ if $k\neq 0$ follows directly from the definition. Assume now that $n\geq 2$. By considering binary vectors $x\in\F_2^n$ as $x=(y,x_{n-1},x_n)$ for $y\in\F_2^{n-2}$, we get:
    \begin{align*}
    \Dnka{n}{k}{a}&=\sum_{x\in \Ekn{k}{n}} (-1)^{d_n(x) +a\cdot x}\\
    &=\sum_{y\in \Ekn{k}{n-2}} (-1)^{d_{n-2}(y) +b\cdot y}+
    \sum_{y\in \Ekn{k-1}{n-2}} (-1)^{d_{n-2}(y) +b\cdot y + a_{n-1}}\\
    &\qquad+\sum_{y\in \Ekn{k-1}{n-2}} (-1)^{d_{n-2}(y) +b\cdot y + a_{n}}
    +\sum_{y\in \Ekn{k-2}{n-2}} (-1)^{d_{n-2}(y) +b\cdot y +1 + a_{n-1}+ a_{n}}\\
    &=\Dnka{n-2}{k}{b}+\left((-1)^{a_{n-1}} + (-1)^{a_{n}}\right)\Dnka{n-2}{k-1}{b} + (-1)^{1+a_{n-1}+a_n} \Dnka{n-2}{k-2}{b}.\qedhere
    \end{align*}
\end{proof}

@@

\begin{remark}\label{rmk:dependence_dnka}
    The recursive formula of Proposition~\ref{prop:recursiveDnka} gives different cases depending on the values of the pair $(a_{n-1},a_n)$. However, the same reasoning applies with any pair of the form $(a_{2i-1}, a_{2i})$ for $i\in[0,n/2]$. Therefore, the value of $\Dnka{n}{k}{a}$ depends only on the number of pairs $(a_{2i-1}, a_{2i})$ in $a$ being $(0,0)$, $(1,1)$, $(0,1)$ or $(1,0)$. We give the values of $\Dnka{2}{k}{a}$ for $n=2$ in Table~\ref{tab:Dnka}. These values together with Proposition~\ref{prop:recursiveDnka} are sufficient to determine any $\Dnka{n}{k}{a}$. We will get back to this in Remark~\ref{rmk:Dnka_dep_pqr}.
\end{remark}

\begin{table}
	\centering
	\begin{tabular}{|c|c|c|c|}
		\hline
		$a$ & $(0,0)$ & $(0,1)$ & $(1,1)$ \\ \hline
		$\Dnka{2}{0}{a}$ &$1$&$1$ & $1$\\ \hline
		$\Dnka{2}{1}{a}$ &$2$& $0$ & $-2$\\ \hline
		$\Dnka{2}{2}{a}$ &$-1$&$1$ & $-1$\\ \hline
	\end{tabular}
	\tablecaption{Values of $\Dnka{2}{k}{a}$ for $n=2$.}\label{tab:Dnka}
\end{table}

In the following, we show how the Walsh transform of $f$ can be written in terms of the $\Dnka{n}{k}{a}$. Then, in Section~\ref{sec:balancedness} we use Proposition~\ref{prop:recursiveDnka} to determine the balancedness of $f$, and in Section~\ref{sec:genfunc} we study the value of $\Dnka{n}{k}{a}$ using generating functions.

\tsnote{@Pierrick, check the below proposition with its proof.}
\begin{proposition}\label{prop:WT}
    Let $n\geq 0$ be an even integer, and let $f\in\BN$ be the revisited \hwbf{}. Denote by $\pi:[1,n]\to[1,n]$ the permutation sending the first $n/2$ elements to the odd positions and the $n/2$ last ones to the even positions. Let $a\in\F_2^n$ be a binary vector, and let $b=\pi(a)+(0,1,0,1,\ldots,0,1)$. Then the following holds true:
	\[ \wt{f}(a)=1+\sum_{k=1}^n \Dnka{n}{k}{b+\pi(e_k)}.\]
\end{proposition}

\begin{proof}
We have by Proposition~\ref{prop:restrWT} that:
\[
\wtk{f}{k}(a)=\sum_{x \in \Ekn{k}{n}} (-1)^{f(x)+ax}=\sum_{x \in \Ekn{k}{n}} (-1)^{\sum_{i=1}^{n/2} x_i x_{i+\frac{n}{2}}+ c \cdot x}=\Dnka{n}{k}{d},\]
where $d$ is the permutation of $c$ obtained by sending the first $n/2$ elements to the odd positions, and the $n/2$ last ones to the even positions.

@@

\subsection{Balancedness of $f$}\label{sec:balancedness}
\tsnote{@Pierrick, check entire section.}

@@

\begin{theorem}\label{th:bal}
	Let $n \in \N^*$ even, and $f$ be the $n$-variable function from Definition~\ref{def:revHWBF}, the following holds on its Walsh transform in $0_n$:
	\[\wt{f}(0_n)= \left \{
	\begin{array}{l l}
	0 & \text{ if } n/2 \text{ is even, } \\
	-2 \binom{n/2-1}{(n-2)/4} & \text{ if }  n/2 \text{ is odd.}
	\end{array}\right.\]

@@

This section examines the values of $\Dnka nka$ through the use of generating functions. To begin, in Section~\ref{sec:genfunc}, we determine essential characteristics of these values. Subsequently, in Section~\ref{sec:Cauchy}, we employ Cauchy's estimate to bound the absolute values of $\Dnka nka$. This approach allows us to also bound the absolute values of the Walsh transform coefficients of the revisited \hwbf{}, thereby constraining the nonlinearity. Lastly, in Section~\ref{sec:general}, we illustrate how this result can be extended to bound the nonlinearity of a family of weightwise quadratic functions.

\subsection{Study of $\Dnka nka$ through generating functions}\label{sec:genfunc}

\begin{definition}\label{defi:P_a}
    Let $n\geq 0$ be an even integer. For $a\in\F_2^n$, we denote by $p=p(a)$ the number of $i\geq 1$ for which $(a_{2i-1},a_{2i})=(0,0)$, by $q=q(a)$ the number of $i\geq 1$ for which $(a_{2i-1},a_{2i})=(1,1)$, and by $r=r(a)$ the number of $i\geq 1$ for which $(a_{2i-1},a_{2i})\in\{(0,1),(1,0)\}$. We then have $p+q+r=n/2$. We further introduce the integer polynomial $P_a(z)$ given by the following expression:
    \[
    P_a(z)=(-z^2+2z+1)^{p}\cdot(-z^2-2z+1)^{q}\cdot(z^2+1)^{r}.
    \]
\end{definition}

\begin{proposition}\label{proposition:generating_fct}
	For every even integer $n\geq 0$ and every binary vector $a\in\F_2^n$, the following holds:
    \[
        \sum_{k\in\Z}\Dnka nka z^k= P_a(z).
    \]
\end{proposition}

\begin{proof}
	We proceed by induction on even $n\geq 0$. 
	For the base case $n=0$, for $a\in\F_2^0$ (the empty vector), we have $P_a(z)=1$, while $\Dnka 00a=1$ and $\Dnka 0ka=0$ if $k\neq 0$ (from Proposition~\ref{prop:recursiveDnka}). 
	For the inductive step, let $n\geq 2$ be even, and assume that the corresponding formula holds for $n-2$. Using the recursive formula for $\Dnka nka$ with the notation $a=(b,a_{n-1},a_n)$, we have:
	\begin{align*}
	\begin{split}
	\sum_{k\in\Z}\Dnka nka z^k&=\sum_{k\in\Z}\Dnka{n-2}kbz^k+\big((-1)^{a_{n-1}}+(-1)^{a_{n}}\big)\sum_{k\in\Z}\Dnka{n-2}{k-1}{b}z^k\\
	&\qquad+(-1)^{1+a_{n-1}+a_n}\sum_{k\in\Z}\Dnka{n-2}{k-2}{b}z^k\\
	&=P_b(z)+\big((-1)^{a_{n-1}}+(-1)^{a_{n}}\big)zP_b(z)+(-1)^{1+a_{n-1}+a_n}z^2P_b(z)\\
	&=P_b(z)\cdot\Big(z^2(-1)^{1+a_{n-1}+a_n}+z\big((-1)^{a_{n-1}}+(-1)^{a_{n}}\big)+1\Big)\\
	&=P_a(z).\qedhere
	\end{split}
	\end{align*}
\end{proof}

\begin{remark}\label{rmk:Dnka_dep_pqr}
    We have already argued in Remark~\ref{rmk:dependence_dnka} that $\Dnka nka$ only depends on $p(a)$, $q(a)$ and $r(a)$. This observation can also be obtained from Proposition~\ref{proposition:generating_fct}: since $P_a(z)$ only depends on $p(a)$, $q(a)$ and $r(a)$ (by definition), we deduce that also $\Dnka nka$ only depends on $p(a)$, $q(a)$ and $r(a)$. Therefore, for any $p,q,r\geq0$ satisfying $p+q+r=n/2$, it makes sense to introduce the notation $\mathsf D_{n,k}^{p,q,r}=\Dnka nka$, where $a\in\mathbb F_2^n$ is any vector with $(p(a),q(a),r(a))=(p,q,r)$. We will make use of this notation later on.
\end{remark}

\begin{proposition}\label{proposition:symmetry}
	For every even integer $n\geq 0$, for every binary vector $a\in\F_2^n$, and for every integer $k$, the following holds:
	\[	\Dnka n{n-k}a=(-1)^{p+q+k}\Dnka nka.	\]
\end{proposition}
